---
title: "Delta Smelt"
author: "Eric Lawrence"
date: "7/19/2021"
output:
  html_document:
    code_folding: show
    code_download: true
    keep_md: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo =TRUE, warning = FALSE, message = FALSE)
```


# Concentration-Response Models

The following concentration-response models were constructed using the data supplied by Hutton et al. (2021) who analyzed the effect of salinity on the toxicity of seven pesticides to silversides (*Menidia beryllina*). The data were available online as part of the supplemental materials.

Data points plotted in the figures have been jittered, or moved around slightly, on x and y axes so that all they are not all stacked on each other.

Hutton, SJ, St. Romain SJ, Pedersen EL, Siddiqui S, Chappell PE, White JW, Armbrust KL, Brander SM. 2021. Salinity alters toxicity of commonly used pesticides in a model euryhaline fish species (*Menidia beryllina*). *Toxics* 9:114.

## drc Model

I used the Log Logistic 3 parameter model included in the drc package to plot the curves. I used the binomial type and fixed the upper limit to 1 (100% mortality).

Log-logistic 3 parameter model:

$f(x) = 0+\frac{d-0}{(1+exp(b(log(x)-e)))}$

where

$d$ = 1 to fix the upper limit to 1

$e$ = LC50

To use this equation in Netica we need to estimate the $b$ and $e$ parameters.

## Salinity

Hutton et al. (2020) used two sets of experiments using 5 PSU and 15 PSU Salinity. I have combined the data from the two salinity values to incorporate uncertainty along a salinity gradient. Hutton et al. (2020) also did not find significant differences between the different salinity models.

## R Code

<details>
  <summary><b>R Code</b></summary>
  
```{r}

library(drc)
library(tidyverse)

smelt <- read.csv("smelt_range_finding_survival_data.csv")


drc.plot <- function (cont) {
  
  ## filter for chemical
  filt <- smelt %>%
  filter(chemical == cont) %>%
  mutate(percent_m = dead / total)
  
  ## Create model
  mod <- drm(filt$percent_m ~ filt$dose.adj, fct = LL.3(fixed = c(NA, 1, NA)), type = "binomial")
  
  ## Plot curve
  
  par(mfrow=c(2,2))
  
  plot(mod, type = "confidence", ylab = "Percent Mortality",
     xlab = "Concentration (mg/L)",
     main = paste(cont, "LL.3"))
  #plot(mod, type = "all", add = TRUE)
  
  y <- jitter(filt$percent_m)
  x <- jitter(filt$dose.adj)
  
  points(y ~ x)


  #Legend
  
  legend("topleft", c("95% Confidence Interval"),
         lty=c(1,1),
         lwd=c(2.5,2.5), 
         col=c('gray'),
         bty = "n",
         cex = 0.8)
  
  ## QA/QC
  
  qqnorm(resid(mod))
  
  
  qqline(resid(mod))
  
  plot(resid(mod)~predict(mod), main = "Residuals")
  abline(h=0)
  
  ############################## Model Parameters

  sum <- summary(mod)
  print(sum)
  
  ################################ EC50

  mod.ec <- ED(mod, c(5, 10, 20, 50), interval = "delta")
  
}

drc.plot.sal <- function (cont) {
  
  ## filter for chemical
  filt <- smelt %>%
  filter(chemical == cont) %>%
  mutate(percent_m = dead / total)
  
  ## Create models
  mod.5 <- drm(filt$percent_m[filt$salinity == 5] ~ filt$dose.adj[filt$salinity == 5], fct = LL.3(fixed = c(NA, 1, NA)), type = "binomial")
  
  mod.15 <- drm(filt$percent_m[filt$salinity == 15] ~ filt$dose.adj[filt$salinity == 15], fct = LL.3(fixed = c(NA, 1, NA)), type = "binomial")
  
  ## Plot curves
  
  plot(mod.5, type = "confidence", ylab = "Percent Mortality",
     xlab = "Concentration (mg/L)",
     main = paste(cont, "LL.3, Salinity"),
     col = "orange")
  #plot(mod.5, type = "all", col = "orange", add = TRUE)
  y.5 <- jitter(filt$percent_m[filt$salinity == 5])
  x.5 <- jitter(filt$dose.adj[filt$salinity == 5])
  points(y.5 ~ x.5, col = "orange")
  
  
  plot(mod.15, type = "confidence", ylab = "Percent Mortality",
     xlab = "Concentration (mg/L)",
     main = paste(cont, "LL.3"),
     col = "blue",
     add = TRUE)
  #plot(mod.15, type = "all", col = "blue", add = TRUE)
  y.15 <- jitter(filt$percent_m[filt$salinity == 15])
  x.15 <- jitter(filt$dose.adj[filt$salinity == 15])
  
  points(y.15 ~ x.15, col = "blue")
  


  #Legend
  
  legend("topleft", c("5 PSU Salinity","15 PSU Salinity"),
         lty=c(1,1),
         lwd=c(2.5,2.5), 
         col=c('orange', "blue"),
         bty = "n",
         cex = 0.8)


  
}

```

</details>
<br><br>

# Bifenthrin

Model Equation

$f(x) = 0+\frac{1-0}{(1+exp(-0.85959(log(x)-0.000013751)))}$


```{r, fig.width=8, fig.height=7}
drc.plot("bifenthrin")
drc.plot.sal("bifenthrin")

```

# Chlorpyrifos

Model Equation

$f(x) = 0+\frac{1-0}{(1+exp(-0.5273(log(x)-0.004397)))}$

```{r, fig.width=8, fig.height=7}

drc.plot("chlorpyrifos")
drc.plot.sal("chlorpyrifos")

```

# Dicloran

Model Equation

$f(x) = 0+\frac{1-0}{(1+exp(-0.6628(log(x)-0.003734)))}$

```{r, fig.width=8, fig.height=7}

drc.plot("dicloran")
drc.plot.sal("dicloran")

```

# Myclobutanil

Model Equation

$f(x) = 0+\frac{1-0}{(1+exp(-0.3286(log(x)-2.315)))}$

```{r, fig.width=8, fig.height=7}

drc.plot("myclobutanil")
drc.plot.sal("myclobutanil")

```

# Paraquat

Model Equation

$f(x) = 0+\frac{1-0}{(1+exp(-0.3334(log(x)-26.1204)))}$

```{r, fig.width=8, fig.height=7}

drc.plot("paraquat")
drc.plot.sal("paraquat")

```

# Penconazole

Model Equation

$f(x) = 0+\frac{1-0}{(1+exp(-0.3363(log(x)-0.04808)))}$

```{r, fig.width=8, fig.height=7}

drc.plot("penconazole")
drc.plot.sal("penconazole")

```

# Triadimefon

Model Equation

$f(x) = 0+\frac{1-0}{(1+exp(-0.6433(log(x)-0.7734)))}$

```{r, fig.width=8, fig.height=7}

drc.plot("triadimefon")
drc.plot.sal("triadimefon")

```

<details>
  <summary><b>R Session Information</b></summary>
```{r results=T}
xfun::session_info('rmarkdown')
```
</details>
<br><br>

